% DSS 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```

## DSS

A small DSS service was implemented for the purpose of predicting the risk using three disease risk prediction models.  
The code provided is an example of a daily model run and will not be fully operational, due to protection of providers of data (Met Eirean) and one of risk prediction models (Blight Management (DK)).  

## Load packages
```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "data.table",
    "ggplot2",
    "knitr",
    "taskscheduleR",
    "zoo",
    "RCurl",
    "imputeTS",
    "scales",
    "padr",
    "mailR",
    "devtools",
    "readxl",
    "miniUI",
    "shiny",
    "htmltools",
    "devtools",
    "stringr",
    "lubridate",
    "readr",
    "cowplot",
    "egg",
    "conflicted"
  )
new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)

#if instal is not working try 
#install.packages("ROCR", repos = c(CRAN="https://cran.r-project.org/"))
```
  
  Only some parts of the script are reproducible. 
  
  The implementation is consisted of the main script which is calling other scripts for: 
1. Pooling the data from ftp  
2. Merging observed and interpolated forecasted values
3. Run models
5. Make graphs
6. Send graphs via email
  
## Main script

```{r main_scrpt,eval=FALSE}
Sys.time()

source("./scripts/lib/Packages.R")

#sync data
source("./scripts/ftp.R")

#Get stations and date
stations <- c(#"Dunsany", 
              "Gurteen", "Oakpark", "Johnstown" , "Moorepark")

time<-  as.Date(Sys.time(), "%Y%B%d", tz="GMT")

# merge obs and forecast for single location  
# infil missing data, and values (see more inside the script)
source("./scripts/Station df.R")
df_list <- lapply(stations, function(x) StationDF(time, x, 
                                                  days_before = 14 #define how many days before we would  like to have on the graph
                                                  ))  #Merge files
names(df_list) <- stations

# Check for missing values and send an email if something is wrong  
source("./scripts/lib/errors/check_missing_values.R")


#run Irish rules model and attach it to each df
source("./scripts/lib/IrishRules.R")
df_list <-lapply(df_list, function(x) mutate(x,irish_rules= IrishRulesModel(x)))

#Run Danish model 
source('./scripts/lib/Danish_Model/Model_BlightManagement.R')
df_list <-lapply(df_list, function(x) mutate(x,dk_daily_spor= Model_BlightManagement(x)[[1]]))
df_list <-lapply(df_list, function(x) mutate(x,dk_infpressure= Model_BlightManagement(x)[[2]]))

#Irish model modified
source("./scripts/lib/IrishRulesModelSensitive.R")
df_list <-lapply(df_list, function(x) mutate(x,irish_rules_sensitive= IrishRulesModelSensitive(x)))


#save outputs every day for analysis 
filenames<-paste0("./data/daily outputs/",as.character(time),"_",names(df_list),".csv")
Map(write.csv,df_list,filenames)


# Create dir for each day
dir.create(paste0("./graphs/daily graphs/",as.character(time)))

# Create and save plots for all stations
source("./scripts/Plot outputs.R")
plot_list <- lapply(df_list, function(x) PlotRisk (x, time))


# Create and save plots for all stations
source("./scripts/Treatment.R")




#Risk status email
source("./scripts/lib/Mailing list.R")
send.mail(
  from = "<cropriskstatus@gmail.com>",
  # to = "<mladencucak@gmail.com>",              #c("<mladencucak@gmail.com>", "recipient2@gmail.com"),
  to = mailing_list,
  subject = paste("Potato late blight risk status for", as.Date(Sys.Date())),
  body = html_body,
  html = TRUE,
  smtp = list(
    host.name = "smtp.gmail.com",
    port = 465,
    user.name = "cropriskstatus",
    passwd = "lateblight",
    ssl = TRUE
  ),
  
  attach.files = list.files(
    paste0("./graphs/daily graphs/", as.character(time)),
    full.names = TRUE,
    pattern = ".png"
  ),
  authenticate = TRUE,
  send = TRUE,
  debug = TRUE
)
Sys.time()
```


  The weather data is provided on an FTP server. 
  
```{r get_data, eval= FALSE}
#Define download function
downloadCSV <- function(filename, folder, baseurl) {
  fileurl <- paste0(baseurl, filename)
  if (!file.exists(str_c(folder, "/", filename))) {
    download.file(fileurl, 
                  destfile = as.character(paste0(folder, "/", filename)),
                  method="libcurl", 
                  mode = "wb")#prevents from adding additional empty rows 
  }
}


userpwd <- "Pass_phrase"

#download files 
url_fore<- paste0("ftp://", userpwd, "@aimsir.ftp.mu.ie/MET/FORECAST/")
local_fore <- "./data/ECMWF"
file_fore <- try(getURL(url_fore, userpwd = userpwd,dirlistonly = TRUE, crlf=TRUE ))
file_fore <- unlist(str_extract_all(file_fore, ".+(.csv)"))

url_obs<- paste0("ftp://", userpwd, "@aimsir.ftp.mu.ie/MET/OBS/")
file_obs <- try(getURL(url_obs, userpwd = userpwd,dirlistonly = TRUE, crlf=TRUE ))
file_obs <- unlist(str_extract_all(file_obs, ".+(.csv)"))
local_obs <- "./data/historical"


lapply(file_fore, function (x) downloadCSV(x, local_fore,url_fore))
lapply(file_obs, function (x) downloadCSV(x, local_obs,url_obs))


#Check if everything is all right
if(length(file_fore)==length(dir(local_fore))){
  message("All forecast files are downloaded!")
} else {
  message("Some forecast files are not downloaded!")
}

if(length(file_obs)==length(dir(local_obs))-1){
  message("All historical files are downloaded!")
} else {
  message("Some historical files are not downloaded!")
}

rm(file_fore, file_obs, local_fore, local_obs, url_fore, url_obs, userpwd, downloadCSV)

```
  
  



  
```{r}
  time <- t
  station <-station_name
  Hrly_date<-time-1
  Hrly_date<- format.Date(Hrly_date, "%Y%b%d")
  
  
  if( station== "Oakpark") {
    station_name <- "OakPark"
  }
  
  # load in observations
  Hrly_df <-  read_csv(paste0("./data/historical/Hrly_Obs_",station_name,"_",Hrly_date,".csv"),
                       col_types = cols(X14 = col_skip(), day = col_skip(),
                                        hour = col_skip(), month = col_skip(), 
                                        station_name = col_skip(), station_no = col_skip(), 
                                        year = col_skip()))
  Hrly_df <- Hrly_df[1:30,]  #get obs unti 5AM on curent day
  Hrly_df$date <- dmy_hms(Hrly_df$date)
  Hrly_df <- Hrly_df[c(1,5,4,7, 2,3,6)]
  
  
  # load in observations from numberof days before
  ObsDailyFun <-  function(days_vec){
    fun_df <- read_csv(paste0("./data/historical/Hrly_Obs_",station_name,"_",format.Date(time-days_vec, "%Y%b%d"),".csv"),
                       col_types = cols(X14 = col_skip(), day = col_skip(),hour = col_skip(), month = col_skip(), 
                                        station_name = col_skip(), station_no = col_skip(), year = col_skip()))
    fun_df <- fun_df[1:24,] 
    fun_df$date <- dmy_hms(fun_df$date)
    fun_df <- fun_df[c(1,5,4,7, 2,3,6)]
    return(fun_df)}
  
  #Define how many days before are to be loaded
  days_before <- 2
  obs_list <- lapply(c(days_before:2), ObsDailyFun)
  Hrly_df_1 <- do.call("rbind", obs_list)
  
  # load in forecast
  ec_date <- format.Date(time, "%Y%m%d")
  ec0103day <- read_csv(paste0("./data/ECMWF/ec0103day",station,"-",ec_date,"00.csv"),skip = 13)
  ec0103day <- ec0103day[6:nrow(ec0103day),]   #take out first five hours of forecast
  ec0305day <- read_csv(paste0("./data/ECMWF/ec0305day",station,"-",ec_date,"00.csv"), skip = 42)
  ec0610day <- read_csv(paste0("./data/ECMWF/ec0610day",station,"-",ec_date,"00.csv"),skip = 32)
  
  #arrange colls for binding
  ec0103day<- (unite(ec0103day, date, 1,2,sep = " ")) 
  ec0103day$date <- as.POSIXct(ec0103day$date, tz = "UTC")
  ec0305day<- (unite(ec0305day, date, 1,2,sep = " ")) 
  ec0305day$date <- as.POSIXct(ec0305day$date, tz = "UTC")
  ec0610day<- (unite(ec0610day, date, 1,2,sep = " ")) 
  ec0610day$date <- as.POSIXct(ec0610day$date, tz = "UTC")
  colnames(ec0305day) <- colnames(ec0103day)
  colnames(ec0610day) <- colnames(ec0103day)
  fore_df <- bind_rows(ec0103day,ec0305day,ec0610day)
  colnames(fore_df) <- tolower(colnames(Hrly_df))
  colnames(fore_df) <- tolower(colnames(Hrly_df_1))
  
  #merge all data together
  df <- bind_rows(Hrly_df_1, Hrly_df, fore_df) 
  
  # #infill missing dates
  df <- padr::pad(df)
  
  infil_gap <- 24
  df1 <- data.frame(df$date)
  df1$temp <- round(na.spline(df$temp, na.rm = FALSE, maxgap = infil_gap),1)
  df1$rh <- round(na.spline(df$rh, na.rm = FALSE, maxgap = infil_gap),0)
  df1$rh  <- sapply(df1$rh, function(x) ifelse(x>100, x<-100, x))
  df1$rainfall <- round(na.replace(df$rainfall, 0),1)
  df1$solar <- round(na.spline(df$solar, na.rm = FALSE, maxgap = infil_gap),1)
  df1$solar <- ifelse(df1$solar < 0, 0, df1$solar)
  
  df1$speed <- round(na.spline(df$speed, na.rm = FALSE, maxgap = infil_gap),0)
  df1$dir <- round(na.spline(df$dir, na.rm = FALSE, maxgap = infil_gap),0)
  names(df1) <- c("date","temp", "rhum","rain","solar", "wdspd", "wddir")
  df1$stna <- station_name
  df1$DayNo <- yday(df1$date)
  
  #Remove last Value because it is single value for that date - models cant run
  df1 <- df1[1:c(length(df1$date)-1),]

```
  
  Run the models. 
  
```{r}
#Run Danish model 
source('./scripts/lib/Danish_Model/Model_BlightManagement.R')
df_list <-lapply(df_list, function(x) mutate(x,dk_daily_spor= Model_BlightManagement(x)[[1]]))
df_list <-lapply(df_list, function(x) mutate(x,dk_infpressure= Model_BlightManagement(x)[[2]]))
```

```{r}
#run Irish rules model and attach it to each df
source("./scripts/lib/IrishRules.R")
df_list <-lapply(df_list, function(x) mutate(x,irish_rules= IrishRulesModel(x)))


#Irish model modified
source("./scripts/lib/IrishRulesModelSensitive.R")
df_list <-lapply(df_list, function(x) mutate(x,irish_rules_sensitive= IrishRulesModelSensitive(x)))

```
  
  
  
```{r}
# Create dir for each day
dir.create(paste0("./graphs/daily graphs/",as.character(time)))

# Create and save plots for all stations
source("./scripts/Plot outputs.R")
plot_list <- lapply(df_list, PlotRisk)

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```
  
  
  
```{r}

```


  
```{r}
session_info()
```
  





























