% Pop

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```

## Popoulation analysis 

 After gaining insight into population structure we can move onto the analysis of 

## Load packages
```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "devtools",
    "here",
    "readxl",
    "poppr",
    "egg"
  )

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)

#if instal is not working try 
#install.packages("package_name", repos = c(CRAN="https://cran.r-project.org/"))
```

## Data import/preparation

```{r data}
fin <-  readRDS(file = here::here("data", "gen data", "final", "gendata.rds") )

```

Prepare the data in genalex format. 

```{r genalex}
fin <- 
  unite(fin, "Ind", idd, year, remove = F)

fin <- 
  unite(fin, "Pop", variety, Genotype, year) 


 popcol <- match("Pop",names(fin))
 idcol <- match("Ind",names(fin))


gen <- 
fin[ , c(idcol,popcol,match("D13",names(fin)):match("X__24",names(fin)))]


path <- here::here("data", "gen data", "final", "pop.csv")

data.frame(ssr = c(12,"PInf_Ireland"), 
           samples = nrow(gen), 
           pop =  length(unique(fin$Pop)))

write.table(data.frame(ssr = 12, samples = nrow(gen), pop =  length(unique(fin$Pop))),
            path,
            sep=",",  col.names=FALSE, row.names = FALSE)

write.table( "PInf_Ireland",
            path,
            sep = ",", row.names = FALSE,
            col.names = FALSE, append = T)
names(gen)[grep("X__", names(gen))] <- ""

write.table(gen,
            path,
            sep = ",", row.names = FALSE,
            col.names = !file.exists("myDF.csv"), append = T)

```

Bruvo charts 
```{r load_data}
path <- here::here("data", "gen data", "final", "pop.csv")

monpop <- read.genalex(path, ploidy = 3)



splitStrata(monpop) <- ~variety/genotype/year

monpop

 monpop$strata<-
  monpop$strata %>% 
mutate(genotype = ifelse( genotype ==  "8A1-1","EU_8_1_A1",
                                   ifelse(genotype ==  "6A1","EU_6_A1",
                                          ifelse(genotype ==  "13A2","EU_13_A2",
                                                 ifelse(genotype ==  "8A1","EU_8_A1","" )))))
 

  
# imsn()

monpop_sub <- popsub(monpop, blacklist = character(0))
monpop_dist <- diss.dist(monpop_sub, percent = FALSE, mat = FALSE)
min_span_net <- poppr.msn(monpop_sub, monpop_dist, showplot = FALSE, include.ties = TRUE)

#Bruvo distance Genotype

setPop(monpop) <- ~genotype
monpop_sub <- popsub(monpop, blacklist = character(0))
min_span_net <- bruvo.msn(monpop_sub, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(1000)

cbbPalette <- c ("#ff33cc",  "#ffff33","#3399ff","#F3DA60") 

```


```{r bruvo1}
min_span_net <- bruvo.msn(monpop, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
plot_poppr_msn(monpop,
               min_span_net,
               mlg = FALSE,
               gadj = 7,
               nodescale = 4,
               nodelab = 1,
               palette =cbbPalette,
               cutoff = 11,
               inds = "none",
               quantiles = FALSE,
               beforecut = TRUE,
               pop.leg = TRUE,
               size.leg = FALSE,
               scale.leg = TRUE,
               layfun = igraph::layout_nicely)




```

```{r bruvoyear}
setPop(monpop) <- ~year
monpop_sub <- popsub(monpop, blacklist = character(0))
min_span_net <- bruvo.msn(monpop_sub, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(1000)

cbbPalette <- c ("#ff33cc",  "#ffff33","#3399ff") 

 
  plot_poppr_msn(
    monpop,
    min_span_net,
    mlg = T,
    gadj = 4,
    inds = "none",
    wscale = TRUE,
    nodescale = 8,
    nodelab = 0.15,
    palette = cbbPalette,
    cutoff = 0.9,
    quantiles = FALSE,
    beforecut = TRUE,
    layfun = igraph::layout_nicely
  )

  
dev.copy(png,filename=here::here("results", "gen","bruvo", "Year.png"),
         width = 1000, height= 700 );
dev.off ()

```


```{r bruvovar}
setPop(monpop) <- ~variety

monpop_sub <- popsub(monpop, blacklist = character(0))
min_span_net <- bruvo.msn(monpop_sub, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(1000)
plot_poppr_msn(monpop,
               min_span_net,
               inds = "none",
               mlg = FALSE,
               gadj = 6,
               nodescale = 8,
               palette = rainbow,
               cutoff = 0.2,
               quantiles = FALSE,
               beforecut = TRUE,
               pop.leg = TRUE,
               size.leg = TRUE,
               scale.leg = TRUE,
               layfun = igraph::layout_nicely)

```


```{r}
setPop(monpop) <- ~year/genotype
monpop_sub <- popsub(monpop, blacklist = character(0))
min_span_net <- bruvo.msn(monpop_sub, replen = c(1, 2, 3), add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(1000)
plot_poppr_msn(monpop,
               min_span_net,
               inds = "none",
               mlg = FALSE,
               gadj = 6,
               nodescale = 8,
               palette = rainbow,
               cutoff = 0.2,
               quantiles = FALSE,
               beforecut = TRUE,
               pop.leg = TRUE,
               size.leg = TRUE,
               scale.leg = TRUE,
               layfun = igraph::layout_nicely)

```


```{r}

set.seed(999)
setPop(monpop) <- ~genotype

#Determine the OS to assign apropriate character string for parallel processing
 parallel_proc <- 
  ifelse(Sys.info()['sysname'] == "Windows", "snow", "multicore") 

 
pramx <- xvalDapc(tab(monpop, NA.method = "mean"), pop(monpop))


#HArd to determine the peek number of PC
set.seed(999)
system.time(pramx <- xvalDapc(tab(monpop, NA.method = "mean"), pop(monpop),
                              n.pca = 10:20, n.rep = 1000,
                              parallel = parallel_proc, 
                              ncpus = 4L))
pramx[-1]
pramx[-1]$`Number of PCs Achieving Highest Mean Success`
#
scatter(pramx$DAPC, 
        # col = other(monpop)$comparePal, 
        cex =2 , 
        legend = TRUE,
        clabel = FALSE, 
        posi.leg = "bottomleft", scree.pca = TRUE,
        posi.pca = "topright", 
        cleg = .9, xax = 1, 
        yax = 2, inset.solid = 1)

dev.copy(png,filename=here::here("results", "gen","dapc", "genotype.png"),
         width = 700, height= 700);
dev.off ()
scatter(pramx$DAPC,1,1, 
        # col=myCol,
        bg="white",
        scree.da=FALSE, legend=TRUE, solid=.4)

```


```{r}

```


```{r}

```

