% DAPC analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```


 Discriminant analysis of principal comonents will give us some more insights into possible clustering within the population.  
 We would like to se eif there are effects of our treatments, variety and fungicide programmes.  
 Another interesting thing to look at is year to year variation of the pathogen population structure. 

## Load packages
```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "devtools",
    "here",
    "readxl",
    "poppr",
    "egg",
    "parallel",
    "conflicted"
  )

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)

#if instal is not working try 
#install.packages("package_name", repos = c(CRAN="https://cran.r-project.org/"))
```

## Data import/preparation

```{r load_data}
path <- here::here("data", "gen data", "final", "pop.csv")

monpop <- read.genalex(path, ploidy = 3)

splitStrata(monpop) <- ~variety/genotype/year

monpop


```

## DAPC
Very good tutorials are avaialble at [adegent page](https://github.com/thibautjombart/adegenet/wiki/Tutorials) and [Grünwald lab web page](https://grunwaldlab.github.io/Population_Genetics_in_R/DAPC.html).  
Seting the paralel processing. This argument should be removed if the code is to be reproduced on machine which can not devote muliple core.   
We will also assign a number of cores to be used. We will set it to be one less than the maximum no. of cores, so we dont 'sufficate' the computer. 

```{r}
#Determine the OS to assign apropriate character string for parallel processing
 parallel_proc <- 
  ifelse(Sys.info()['sysname'] == "Windows", "snow", "multicore") 

cpus <-  ifelse(detectCores()>1, c(detectCores()-1), 1)
cpus
```


```{r dapc_var, out.width='100%', fig.width = 6, fig.height=6}
setPop(monpop) <- ~variety


set.seed(999)
system.time(pramx_var <- xvalDapc(tab(monpop, NA.method = "mean"), pop(monpop),
                              n.pca = 5:20, n.rep = 3000,
                              parallel = parallel_proc, 
                              ncpus = cpus))

#Reorder levels of the factor for plotting
pramx_var$DAPC$grp <- 
  factor(pramx_var$DAPC$grp, levels = c("KE", "BQ", "RO", "SE", "CL", "SM"))

saveRDS(pramx_var, file = here::here("results", "gen","dapc",  "dapc_var.RDS"))

pramx_var[-1]
pramx_var[-1]$`Number of PCs Achieving Highest Mean Success`




scatter(
  pramx_var$DAPC,
  legend = TRUE,
  cex = 3,
  # pch=18:23,
  posi.leg = "topleft",
  scree.da = FALSE,
  clab =2,
  cleg = 2.1,
  cell = 1, #zoom
  xax = 1,
  yax = 2,
  inset.solid = 1,
  cstar=0.5,
  lwd=.1
)



```
   
## Treatment effect
  Prepare the data in genalex format once again. We can define only thre stratums, so we will prepare another file which will be re-used for same purpose as above.  

```{r genalex_new}
fin <-  readRDS(file = here::here("data", "gen data", "final", "gendata.rds") )

fin <- 
  unite(fin, "Ind", idd, year, remove = F)

fin <- 
  unite(fin, "Pop", treatment, Genotype, year) 


 popcol <- match("Pop",names(fin))
 idcol <- match("Ind",names(fin))


gen <- 
fin[ , c(idcol,popcol,match("D13",names(fin)):match("X__24",names(fin)))]


path <- here::here("data", "gen data", "final", "pop2.csv")

data.frame(ssr = c(12,"PInf_Ireland"), 
           samples = nrow(gen), 
           pop =  length(unique(fin$Pop)))

write.table(data.frame(ssr = 12, samples = nrow(gen), pop =  length(unique(fin$Pop))),
            path,
            sep=",",  col.names=FALSE, row.names = FALSE)

write.table( "PInf_Ireland",
            path,
            sep = ",", row.names = FALSE,
            col.names = FALSE, append = T)
names(gen)[grep("X__", names(gen))] <- ""

write.table(gen,
            path,
            sep = ",", 
            row.names = FALSE,
            col.names = !file.exists("myDF.csv"), append = T)

```
  
  So now we have created another file in genalex format, pop2.csv.

```{r import_genalex}
path <- here::here("data", "gen data", "final", "pop2.csv")

monpop <- read.genalex(path, ploidy = 3)

splitStrata(monpop) <- ~treatment/genotype/year

monpop
```



```{r dapc_trt, out.width='100%', fig.width = 6, fig.height=6}
setPop(monpop) <- ~treatment

set.seed(999)
system.time(pramx_trt <- xvalDapc(tab(monpop, NA.method = "mean"), pop(monpop),
                              n.pca = 5:20, 
                              n.rep = 3000,
                              parallel = parallel_proc, 
                              ncpus = cpus))

pramx_trt$DAPC$grp <- 
  factor(pramx_trt$DAPC$grp, levels = c("0", "100", "50", "IR", "BM", "MIR"))

saveRDS(pramx_trt, file = here::here("results", "gen","dapc",  "dapc_trt.RDS"))

pramx_trt[-1]

scatter(
  pramx_trt$DAPC,
  legend = TRUE,
  cex = 3,
  # pch=18:23,
  posi.leg = "topleft",
  scree.da = FALSE,
  clab =2,
  cleg = 2.1,
  cell = 1, #zoom
  xax = 1,
  yax = 2,
  inset.solid = 1,
  cstar=0.5,
  lwd=.1
)


```

## Year effect

```{r dapc_year, out.width='100%', fig.width = 6, fig.height=6}
setPop(monpop) <- ~year

set.seed(999)
system.time(pramx_yr <- xvalDapc(tab(monpop, NA.method = "mean"), pop(monpop),
                              n.pca = 5:20,
                              n.rep = 3000,
                              parallel = parallel_proc, 
                              ncpus = cpus))
pramx_yr$DAPC$grp <- 
  factor(pramx_yr$DAPC$grp, levels = c("2016", "2017", "2018", "2019"))

saveRDS(pramx_yr, file = here::here("results", "gen","dapc",  "dapc_yr.RDS"))

pramx_yr[-1]


scatter(
  pramx_yr$DAPC,
  legend = TRUE,
  cex = 3,
  # pch=18:23,
  posi.leg = "topleft",
  scree.da = FALSE,
  clab =1.8,
  cleg = 2,
  cell = 1, #zoom
  xax = 1,
  yax = 2,
  inset.solid = 1,
  cstar=0.5,
  lwd=.1
)

```




```{r}

```


```{r}
session_info()
```

