%Wth

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```

```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "devtools",
    "egg",
    "ggpubr",
    "zoo",
    "kableExtra",
    "conflicted",
    "here",
    "lubridate"
  )

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)
conflict_prefer("year", "lubridate")
conflict_prefer("filter", "dplyr")

#if instal is not working try 
#install.packages("ROCR", repos = c(CRAN="https://cran.r-project.org/"))
```


