%Wth

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```

```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "devtools",
    "egg",
    "ggpubr",
    "zoo",
    "kableExtra",
    "conflicted",
    "here",
    "lubridate"
  )

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)
conflict_prefer("year", "lubridate")
conflict_prefer("filter", "dplyr")

#if instal is not working try 
#install.packages("ROCR", repos = c(CRAN="https://cran.r-project.org/"))
```


## Load the data
 
```{r cars}
df <- 
read_csv(
  here::here("data", "weather" , "hly375.csv"),
  col_types = cols(
    clamt = col_skip(),
    clht = col_skip(),
    vis = col_skip(),
    w = col_skip(),
    ww = col_skip()
  ),
  skip = 17
)

df$date <-
  lubridate::dmy_hm(df$date)
df<- add_column(df, short_date = as.Date(df$date), .after=1)
df<- add_column(df, year = year(df$date), .after=2)
df<- add_column(df, month = month(df$date), .after=3)
df<- add_column(df, day = day(df$date), .after=4)


infil_gap <- 8 #Maximum length of the infill gap
df$temp <-
  round(na.spline(df$temp, na.rm = FALSE, maxgap = infil_gap), 1)
df$rhum <-
  round(na.spline(df$rhum, na.rm = FALSE, maxgap = infil_gap), 0)
df$rhum  <- sapply(df$rhum, function(x)
  ifelse(x > 100, x <- 100, x))

df <- 
df %>% 
  filter(year %in% 2016:2019)%>%
  filter(month %in% c(4:9)) 

#Check if the imputation worked
df %>%
  summarize(
    NA_rain = sum(is.na(rain)),
    NA_temp = sum(is.na(temp)),
    NA_rhum = sum(is.na(rhum))
  )%>% 
  kable(format = "html") %>% 
  kableExtra::kable_styling(latex_options = "striped",full_width = FALSE)


#Only two missing values for rain should not be a problem, so we will just do a linear interpolation
df$rain <- 
  round(na.approx(df$rain, na.rm = FALSE, maxgap = infil_gap), 1)


df %>%
  group_by(year, month, short_date) %>% 
  summarize(temp = round(mean(temp),1),
            rain = sum(rain),
            rhum = round(mean(rhum),1)
            
  ) %>%
  mutate(rain = ifelse(rain == 0, NA, rain)) %>% 
  rename(date = short_date) %>% 
  ggplot() +
  # geom_hline(aes())
  geom_line(
    aes(
      x = date,
      y = rhum,
      color = "Average relative humidity (%)",
      fill = "Average relative humidity (%)"
    ),
    size = 0.25
  ) +
  geom_line(aes(
    x = date,
    y = temp,
    colour = "Average temperature (˚C)",
    fill = "Average temperature (˚C)",
  ),
  size = 0.25) +
  geom_col(
    aes(date,
        rain,
        colour = "Total precipitation (mm/day)",
        fill = "Total precipitation (mm/day)"),
    size = .8 ,
    inherit.aes = TRUE,
    width = 0.03
  ) +
  scale_colour_manual(
    "Daily weather:",
    values = c(
      "Average relative humidity (%)" = "#0EBFE9",
      "Average temperature (˚C)" = "#ED2939",
      "Total precipitation (mm/day)" = "blue"
    )
  ) +
  scale_fill_manual(
    "Daily weather:",
    values = c(
      "Average relative humidity (%)" =  "#0EBFE9",
      "Average temperature (˚C)" = "#ED2939",
      "Total precipitation (mm/day)" = "blue"
    )
  ) +
  scale_linetype_manual(name = "Risk Limits",
                        values = c(1, 2, 3)) +
  scale_y_continuous( breaks = seq(0,100,20))+
  # theme_article() +
  scale_x_date(date_labels = "%b", date_breaks = "1 month")+
  facet_wrap( ~ year, scales = "free", ncol = 1,strip.position="right") +
  guides(fill=guide_legend(nrow=1,byrow=TRUE,title.position = "top"),
         color=guide_legend(nrow=1,byrow=TRUE,title.position = "top"))+
    theme_bw()+
  theme(
    text = element_text(size=13),
    # axis.text.x = element_blank(),
    # axis.ticks.x = element_blank(),
     axis.title.y = element_blank(),
     axis.title.x = element_blank(),
    # axis.text.y = element_blank(),
    # axis.ticks.y = element_blank(),
    legend.position = "top",
    strip.background = element_blank(), 
    strip.placement = "outside"
    # panel.grid.minor =   element_blank(),
        # panel.grid.major =   element_line(colour = "white",size=0.75)
  ) +

ggsave(filename= here::here("results", "wth", "all_wth.png"), 
       width = 7, height =6.5, dpi = 620)

```
