%Wth

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)

chooseCRANmirror(graphics=FALSE, ind=1)
knitr::knit_hooks$set(
  source = function(x, options) {
    hook.r = function(x, options) {
      fence <- "```"
      language = tolower(options$engine)
      if (language == 'node') language = 'javascript'
      if (!options$highlight) language = 'text'
      if(!is.null(options$foldcode)) {
      paste0('\n\n', "<details><summary>The code</summary>\n", fence, language, '\n', x, fence,  '\n\n', "</details>\n")
      } else {
              paste0('\n\n', fence, language, '\n', x, fence,  '\n\n')
      }
    }
    x = knitr:::hilight_source(x, 'markdown', options)
    hook.r(
      paste(c(
        x, 
        ''
      ), collapse = '\n'), 
      options
    )
  }
)

```

```{r libs, warning = FALSE, message= FALSE}
list.of.packages <-
  c(
    "tidyverse",
    "devtools",
    "egg",
    "ggpubr",
    "zoo",
    "kableExtra",
    "conflicted",
    "here",
    "lubridate"
  )

new.packages <-
  list.of.packages[!(list.of.packages %in% installed.packages()[, "Package"])]

#Download packages that are not already present in the library
if (length(new.packages))
  install.packages(new.packages)

packages_load <-
  lapply(list.of.packages, require, character.only = TRUE)

#Print warning if there is a problem with installing/loading some of packages
if (any(as.numeric(packages_load) == 0)) {
  warning(paste("Package/s: ", paste(list.of.packages[packages_load != TRUE], sep = ", "), "not loaded!"))
} else {
  print("All packages were successfully loaded.")
}
rm(list.of.packages, new.packages, packages_load)
conflict_prefer("year", "lubridate")
conflict_prefer("filter", "dplyr")

#if instal is not working try 
#install.packages("ROCR", repos = c(CRAN="https://cran.r-project.org/"))
```
```{r}
## Load the data
 df <- 
  read_csv(
    here::here("data", "weather" , "hly375.csv"),
    col_types = cols(
      clamt = col_skip(),
      clht = col_skip(),
      vis = col_skip(),
      w = col_skip(),
      ww = col_skip()
    ),
    skip = 17
  )

df$date <-
  lubridate::dmy_hm(df$date)
df<- add_column(df, short_date = as.Date(df$date), .after=1)
df<- add_column(df, year = year(df$date), .after=2)
df<- add_column(df, month = month(df$date), .after=3)
df<- add_column(df, day = day(df$date), .after=4)


infil_gap <- 8 #Maximum length of the infill gap
df$temp <-
  round(na.spline(df$temp, na.rm = FALSE, maxgap = infil_gap), 1)
df$rhum <-
  round(na.spline(df$rhum, na.rm = FALSE, maxgap = infil_gap), 0)
df$rhum  <- sapply(df$rhum, function(x)
  ifelse(x > 100, x <- 100, x))

df <- 
  df %>% 
  filter(year %in% 2016:2019)%>%
  filter(month %in% c(4:9)) 

#Check if the imputation worked
df %>%
  summarize(
    NA_rain = sum(is.na(rain)),
    NA_temp = sum(is.na(temp)),
    NA_rhum = sum(is.na(rhum))
  )%>% 
  kable(format = "html") %>% 
  kableExtra::kable_styling(latex_options = "striped",full_width = FALSE)


#Only two missing values for rain should not be a problem, so we will just do a linear interpolation
df$rain <- 
  round(na.approx(df$rain, na.rm = FALSE, maxgap = infil_gap), 1)





dfday <-
  df %>%
  group_by(year, month, short_date) %>%
  summarize(
    mintemp = round(min(temp), 1),
    minrhum = round(min(rhum), 1),
    temp = round(mean(temp), 1),
    rain = sum(rain),
    rhum = round(mean(rhum), 1),
    meantemp = round(mean(temp), 1),
    meanrhum = round(mean(rhum), 1)
  )

#REad disease data
disdata <- 
read_rds( path = here::here("results", "dis", "Duration of the epidemics in years.Rdata")) 
colnames(disdata) <- 
  c("year", "disout", "disend")
dfday <-
  left_join(dfday, disdata, by = "year")

# rolling mean
rlmean <- 10
dfday$wrhum <- zoo::rollmean(dfday$meanrhum, rlmean, mean, align = 'center', fill = NA)
dfday$wtemp <- zoo::rollmean(dfday$mintemp, rlmean, mean, align = 'center', fill = NA)

#conducive days
templine <- 10
dfday <- 
  dfday %>% 
  mutate(condday = ifelse( wtemp>=templine&rain>=.2&rhum>=80& short_date<disout, 50,NA))

#sum of conditions prior to outbreak
dfday <- 
dfday %>% 
  group_by(year) %>% 
  summarise(consum= sum(condday/50, na.rm = TRUE )) %>%
  left_join(dfday,., by = "year")

#LAbels for cumulative sum of critical days
dfday <- 
dfday %>% 
  mutate(symptoms = paste0("Symptoms (", consum, " cond. days)"))


linesize <- .35 #size of the line for variables

dfday %>% 
  mutate(rain = ifelse(rain == 0, NA, rain)) %>% 
  rename(date = short_date) %>% 
  ggplot() +
  geom_point(aes(disout, 46),shape =2, size = 1.5, fill ="black")+
  geom_point(aes(disend, 46),shape =2, size = 1.5, fill ="black")+
  geom_point(aes(date, condday), shape = 3,size = .55)+
  geom_line(
    aes(
      x = date,
      y = rhum,
      colour = "Relative humidity (%)"
    ),
    size = linesize
  ) +
  geom_line(aes(
    x = date,
    y = temp,
    colour = "Temperature (˚C)",
  ),
  size = linesize) +
  geom_line(aes(
    x = date,
    y = wtemp,
    colour = "Rolling mean temp. (˚C)"
  ),
  linetype = "dashed",
  size = linesize) +
  geom_col(
    aes(date,
        rain,
        fill = "Total precipitation (mm/day)"),
    size = 1.2 ,
    inherit.aes = TRUE,
    width = 0.8
  ) +
  scale_colour_manual(
    "Daily weather:",
    values = c(
      "Relative humidity (%)" = "#0EBFE9",
      "Temperature (˚C)" = "#ED2939",
      "Rolling mean temp. (˚C)"= "darkred"
    )
  ) +
  scale_fill_manual(name= NA, values = c("Total precipitation (mm/day)" = "blue")
  ) +
  scale_y_continuous( breaks = seq(0,100,10), limits = c(-1,100),expand = c(0, 0))+
  scale_x_date(date_labels = "%b", date_breaks = "1 month",expand = c(.03, .04))+
  facet_wrap( ~ year, scales = "free", ncol = 1,strip.position="right") +
  guides(fill=guide_legend(nrow=1,byrow=TRUE,title.position = "top"),
         color=guide_legend(nrow=1,byrow=TRUE,title.position = "top"),
         linetype=guide_legend(nrow=1,byrow=TRUE,title.position = "top"))+
  theme_bw()+
  theme(
    text = element_text(size=10.8),
    # axis.text.x = element_blank(),
    # axis.ticks.x = element_blank(),
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    # axis.text.y = element_blank(),
    # axis.ticks.y = element_blank(),
    legend.position = "top",
    strip.background =element_rect(colour = "black",size=0.75), 
    strip.placement = "outside",
    legend.title = element_blank(),
    panel.grid.minor =   element_blank(),
    panel.grid.major =   element_line(colour = "lightgray",size=linesize),
    plot.margin = margin(0, 0, 0, 0, "cm"),
    panel.spacing.y=unit(0, "lines"),
    legend.box.spacing = unit(0.1, 'cm'),
    legend.margin = margin(0, 0, 0, 0, "cm"),
  )+
  geom_text(aes(label= symptoms, y = 39, x = disout),size = 2.5)+
  geom_text(aes(label= "Last assesment", y = 39, x = disend-1.5),size = 2.5)+
  ggsave(filename= here::here("results", "wth", "all_wth_dis.png"),
         width = 6.5, height =8, dpi = 720)

# shell.exec( here::here("results", "wth", "all_wth_dis.png"))





```


